github link (https://github.com/panchenying/test1.git)

#HYPOTHESIS
I will be answering the question: What childhood (8-9 years) risk factors predict juvenile (10-16 years) offending?
Based on this a priori hypothesis from readings I've done, I will be looking at variables related to antisocial child behaviour, low intelligence and school acheivement, family criminality, family poverty, and poor parenting and neglect.

loading multiple packages
```{r}
library(pacman)
p_load(dplyr, ggplot2, magrittr, psych, tidyverse, haven, broom, purrr)
```
loading data
```{r}
conviction <- read_spss("conviction_data.sav")
delinquency <- read_spss("Cambridge_delinquency.sav")
```
creating a smaller dataset by selecting only the variables I want to look at
```{r}
juvenile_delinquency <- delinquency %>%  select(v34, v39, v42, v53, v108, v103, v147,
                               v158, v161, v137, v300, v301, v169, v170, v171, v172, v175, v177,
                               v83, v84, v92, v138, v89, v90, v59, v60, v62, v63, v123)
```



